### 2022-02-13 chapters/conclusions/conclusions.tex
commentaires du patron

\section{Conclusions}
In this thesis, we have developed a framework to better understand the propagation and onset of the SCTLD, as well as the impact of major hurricanes onctransport processes over coral reefs. This model had to be able to accurately capture ocean circulation at different scales to study transport processes in the FCR. First, the large-scale Loop Current-Florida Current system had to be accurately reproduced to correctly capture its impact on eddy formation near the DRTO and the Florida Keys. Second, the model had do reproduce ocean circulation at the reef-scale to capture recirculation and acceleration of currents between reefs and islands. This was performed by coupling the multi-scale ocean model SLIM to a connectivity-based epidemiological model (chapter \ref{chap:sctld}), a sediment transport model (chapter \ref{chap:onset}) and a spectral wave model (chapter~\ref{chap:irma}).

Our coupled hydro-epidemiological model reproduced the observed propagation of the SCTLD through the FRT by modeling the transport of its causative agent within neutrally buoyant material driven by mean barotropic currents. It confirmed the result of previous ex-situ studies that showed evidence of waterborne transmission of the SCTLD with an average transmission time of the order of 10 days. Furthermore, by using coral resistance to the disease as a parameter, our model results suggested that, on average, corals had a rather low resistance against the disease. This confirmation of experimental results by model results illustrates the important potential of models in evaluating hypotheses derived from field observations on large scale systems. As such, models informed and confronted against field knowledge are a powerful tool to inform the management of complex ecosystems such as the FCR.

This epidemiological model was then used to build networks of potential exchanges of disease agents between $\sim 1.6\times10^4$ sub-reef areas during a period of three years. The analysis of these networks showed that exchanges from the Marquesas to the DRTO were interrupted during most of 2020. This interruption was likely linked to eddy activity near the DRTO and was consistent with the apparent stalling of the outbreak in the region in 2020. This confirms the role played the Loop Current/Florida Current system in the control of the connectivity in the Florida Keys and the DRTO. Furthermore, sub-reefs predicted to be more vulnerable to SCTLD in the disease networks were consistent with the observed geographic distribution of disease coral in the DRTO. This further highlights that modeled hydrodynamics are highly explanatory of the propagation of the disease.

The results of chapters \ref{chap:sctld} and \ref{chap:drto} show that our high-resolution biophysical is a valuable tool to understand the spread of SCTLD in Florida. The model successfully reproduced the observed propagation of the disease in 2018-2019 and allowed us to link the observed stalling of SCTLD to observable hydrodynamic features. Furthermore, the model allowed us to deduce potential characteristics of the disease agent and its vector. This highlights the potential of  model to understand the driving mechanics of ongoing marine disease outbreaks. Furthermore, indicators such as the vulnerablity and source indicators developed in chaper \ref{chap:drto} could be useful to predict and/or mitigate the spread of SCTLD in other territories of the Caribbean.

@@

Finally, we developed a couple wave-current model to study the impact of Hurricane Irma on transport processes in the Florida Keys. This coupled model correctly reproduced the observed wave and hydrodynamics during the passage of the hurricane. Furthermore, model results suggest that an import dissipation occurred over coral reefs through depth-induced wave-breaking and bottom friction. This important dissipation produced a large wave radiation stress gradients, which caused differences in modeled velocity of up to 1 m/s between the coupled and uncoupled model runs. However, wave-induced transport was dominated by the Stokes drift, which yielded an impact four times larger than the radiation stress gradient. Results of chapter \ref{chap:irma} strongly advocates for the use of coupled wave-currents to accurately model the dispersal of coral larvae under storm conditions. Although we built the tools to study the impact of hurricane on the exchanges of larvae between reefs by developing the coupled SLIM+SWAN, we did not evaluate the impact of Irma on coral connectivity by lack of time.

@@

\subsection*{Combining disease and larval connectivity}
All chapters dedicated to SCTLD rely on the concept of connectivity. In this context, connectivity negatively impacts coral reefs as it drives the propagation of disease agents. However, connectivity acts as a double-edged sword. Larval connectivity promotes the resilience of the reef network as it allows the recolonization of depopulated reefs after extreme events such as mass bleaching, hurricanes or disease outbreaks. Larval connectivity is therefore a valuable tool to inform coral reef protection and restoration. Identifying reefs with large out-degree our centrality in the connectivity network indicates the optimal sites for coral outplants in the context of active restoration. Furthermore, knowledge about larval connectivity would allow us to build robust and sustainable networks of connected marine protected areas. As SLIM allows the computation of larval connectivity \citep{thomas2014numerical,frys2020fine}, a possible application of the tools developed in this thesis is the combination of both types of connectivity.  This combined information would allow us to identify reefs that have high potential for restoration and low vulnerability to the disease. A possible measure of the restoration potential is the (weighted) out-degree of reefs in the larval connectivity network, as reefs with strong and numerous outgoing connections in the network are more likely to send many larvae to many connected reefs. Vulnerability to disease can be evaluated using the (weighted) in-degree of reefs in the disease connectivity network, as reefs with strong and numerous incoming connections are more likely to receive more disease agents from many connected reefs. Therefore, reefs with high restoration potential and low vulnerability would correspond to reefs with large larval(weighted) out-degree and low disease (weighted) in-degree. This approach have been applied in \cite{holstein2022} and could be further developed in future studies using more complex connectivity metrics such as the PageRank index \citep{frys2020fine}.

\subsection*{Improving the epidemiological model}
A limitation of the epidemiological model developed in chapter \ref{chap:sctld} is that it uses species-averaged parameters and assumes an uniform distribution of coral reefs throughout the FRT. The fact that our model correctly reproduced the observed spread of SCTLD for a well-defined range of values of coral resistance to the disease highlights the important impact of this parameter on the propagation of the disease. However, coral susceptibility to SCTLD can significantly vary between species and ranges from highly susceptible (\eg~\textit{Dichocoenia stokesii}, \textit{Meandrina meandrites}) to intermediately susceptible (\eg~\textit{Orbicella faveolata}, \textit{Montastrea cavernosa}) and weakly susceptible (\eg~\textit{Acropora Palmata}, \textit{Acropora cervicornis}). A first improvement could therefore be to consider classes od corals with different susceptibilities and hence use different infection and removal parameter values in the model. Such improvement would be technically challenging as it would increase the number of variables of model. The development of the model could therefore require the use of more advance graph theory tools such as the multilayer networks \citep{kivela2014multilayer,pilosof2017multilayer}. Furthermore, increasing the number of parameters will thus make their calibration more complex. Moreover, this new modeling framework would require estimates of the distribution of the different coral species across the FRT. Again, compiling this data and deriving realistic estimates would represent a significant challenge. Adding species-specific or location-specific parameters in the model would also allow the evaluation of the impact of disease mitigating actions. Lesion treatment using antibiotics could for instance be taken into account in the model by locally reducing the transmission and/or removal parameters. The model parameters could also be made time-dependent as SCTLD progression was reported to slow down during warmer months in the US Virgin Islands \citep{meiling20203d}. Another improvement would be the inclusion of additional processes in the model. For instance, reinfection by SCTLD is currently not possible as corals are "removed" once they are affected by the disease. Given the high mortality of the disease, the "removed" state of the model was assumed equivalent to "dead". However, one might want to discriminate between coral recovery and mortality. Yet, again, adding processes implies an increase in the number of parameters to be calibrated. This modifications could lead to an endemic disease dynamics.

@@

% Last nice chapter
The work presented in this thesis has demonstrated the capabilities of modeling tools to inform the management of Florida's Coral Reef. Models allow the evaluation of hypotheses derived from field observations on large-scale systems and can help to deduce the characteristics and underlying mechanisms of complex processes. Furthermore, they can be used in a predictive manner to evaluate the impacts of changing environmental conditions on a given ecosystem. Additionally, the framework developed in this thesis relies on the same tools as already widely used larval transport models and could be combined with them to optimize the design of coral reef protection and restoration campaigns. Given this high potential, further research to improve the model and synergistic work with field an lab studies may well yield results which could help maintaining the biological functions of coral reefs as long as possible through the Anthropocene.

>>>>

\section{Conclusions}
In this thesis, we have developed a modeling framework to better understand the propagation and onset of the SCTLD, as well as the impact of major hurricanes on transport processes over coral reefs. Our model had to be able to accurately capture ocean circulation patterns at different scales to study transport processes in the FCR. First, the large-scale Loop Current-Florida Current system had to be accurately reproduced to correctly capture its impact on eddy formation near the DRTO and the Florida Keys. Second, the model had to reproduce ocean circulation at the reef-scale to capture the recirculation and acceleration of currents between reefs and islands. This was performed by coupling the multi-scale ocean model SLIM to a connectivity-based epidemiological model (chapter \ref{chap:sctld}), a sediment transport model (chapter \ref{chap:onset}) and a spectral wave model (chapter~\ref{chap:irma}).

Our coupled hydro-epidemiological model reproduced the observed propagation of the SCTLD through the FRT by representing the transport of its causative agent within neutrally buoyant material driven by mean barotropic currents. It confirmed the result of previous ex-situ studies that showed evidence of waterborne transmission of the SCTLD with an average transmission time of the order of 10 days. Furthermore, by using coral resistance to the disease as a parameter, our model results suggested that, on average, corals had a rather low resistance against the disease. This confirmation of experimental results by model results illustrates the important potential of models in evaluating hypotheses derived from field observations on large scale systems. As such, models informed by and confronted against field knowledge are a powerful tool to inform the management of complex ecosystems such as the FCR.

This epidemiological model was then used to build networks of potential exchanges of disease agents between $\sim 1.6\times10^4$ sub-reef areas composing the FRT during a period of three years. The analysis of these networks showed that exchanges from the Marquesas to the DRTO were interrupted during most of 2020. This interruption was likely linked to eddy activity near the DRTO and was consistent with the apparent stalling of the outbreak in the region in 2020. This confirms the role played the Loop Current/Florida Current system in the control of the connectivity pathways in the Florida Keys and the DRTO. Furthermore, sub-reefs predicted to be more vulnerable to SCTLD in the disease networks were consistent with the observed geographic distribution of diseased corals in the DRTO. This further highlights that the modeled hydrodynamics is highly explanatory of the propagation of the disease.

The results of chapters \ref{chap:sctld} and \ref{chap:drto} show that our high-resolution biophysical model is a valuable tool to understand the spread of SCTLD in Florida. The model successfully reproduced the observed propagation of the disease in 2018-2019 and allowed us to link the observed stalling of SCTLD to observable hydrodynamic features. Furthermore, the model allowed us to deduce potential characteristics of the disease agent and its vector. This highlights the potential of biophysical models to understand the driving mechanics of ongoing marine disease outbreaks. Furthermore, indicators such as the vulnerablity and source indicators developed in chaper \ref{chap:drto} could be useful to predict and/or mitigate the spread of SCTLD in other territories of the Caribbean.

@@

Finally, we developed a couple wave-current model to study the impact of Hurricane Irma on transport processes in the Florida Keys. This coupled model correctly reproduced the observed waves and current dynamics during the passage of the hurricane. Furthermore, model results suggest that an important dissipation occurred over coral reefs through depth-induced wave-breaking and bottom friction. This important dissipation produced a large wave radiation stress gradients, which caused differences in modeled velocity of up to 1 m/s between the coupled and uncoupled model runs. However, wave-induced transport was dominated by the Stokes drift, which yielded an impact four times larger than the radiation stress gradient. Results of chapter \ref{chap:irma} strongly advocates for the use of coupled wave-currents to accurately model the dispersal of coral larvae under storm conditions. Although we built the tools to study the impact of hurricane on the exchanges of larvae between reefs by developing the coupled SLIM+SWAN, we did not evaluate the impact of Irma on coral connectivity by lack of time.

@@

\subsection*{Combining disease and larval connectivity}
All chapters dedicated to SCTLD rely on the concept of connectivity. In this context, connectivity negatively impacts coral reefs as it drives the propagation of disease agents. However, connectivity is actually a double-edged sword. Larval connectivity promotes the resilience of the reef network as it allows the recolonization of disturbed reefs following extreme events such as mass bleaching, hurricanes or disease outbreaks. Larval connectivity is therefore a valuable tool to inform coral reef protection and restoration. Identifying reefs that are both good sources and sinks of larvae the connectivity network indicates the optimal sites for coral outplants in the context of active restoration. Furthermore, knowledge about larval connectivity would allow us to build robust and sustainable networks of connected marine protected areas. As SLIM allows the computation of larval connectivity \citep{thomas2014numerical,frys2020fine}, a possible application of the tools developed in this thesis is the combination of both types of connectivity.  This combined information would allow us to identify reefs that are good larval sources and have low vulnerability to the disease. A possible measure of larval supply to the network is the (weighted) out-degree of reefs in the larval connectivity network, as reefs with strong and numerous outgoing connections in the network are more likely to send many larvae to many connected reefs. Vulnerability to disease can be evaluated using the (weighted) in-degree of reefs in the disease connectivity network, as reefs with strong and numerous incoming connections are more likely to receive more disease agents from many connected reefs. Therefore, reefs with high restoration potential would correspond to reefs with large larval(weighted) out-degree and low disease (weighted) in-degree. This approach have been applied in \cite{holstein2022} and could be further developed in future studies using more complex connectivity metrics such as the PageRank index \citep{frys2020fine}.

\subsection*{Improving the epidemiological model}
A limitation of the epidemiological model developed in chapter \ref{chap:sctld} is that it uses species-averaged parameters and assumes an uniform distribution of coral species throughout the FRT. The fact that our model correctly reproduced the observed spread of SCTLD for a well-defined range of values of coral resistance to the disease highlights the important impact of this parameter on the propagation of the disease. However, coral susceptibility to SCTLD can significantly vary between species and ranges from highly susceptible (\eg~\textit{Dichocoenia stokesii}, \textit{Meandrina meandrites}) to intermediately susceptible (\eg~\textit{Orbicella faveolata}, \textit{Montastrea cavernosa}) and weakly susceptible (\eg~\textit{Acropora Palmata}, \textit{Acropora cervicornis}). A first improvement could therefore be to consider classes of corals with different susceptibilities and hence use different infection and removal parameter values in the model. Such improvement would be technically challenging as it would increase the number of variables of model. The development of the model could therefore require the use of more advance graph theory tools such as the multilayer networks \citep{kivela2014multilayer,pilosof2017multilayer}. Furthermore, increasing the number of parameters will thus make their calibration more complex. Moreover, this new modeling framework would require estimates of the distribution of the different coral species across the FRT. Again, compiling this data and deriving realistic estimates would represent a significant challenge. Adding species-specific or location-specific parameters in the model would also allow the evaluation of the impact of disease mitigating actions. Lesion treatment using antibiotics could for instance be taken into account in the model by locally reducing the transmission and/or removal parameters. The model parameters could also be made time-dependent as SCTLD progression was reported to slow down during warmer months in the US Virgin Islands \citep{meiling20203d}. Another improvement would be the inclusion of additional processes in the model. For instance, reinfection by SCTLD is currently not possible as corals are "removed" once they are affected by the disease. Given the high mortality of the disease, the "removed" state of the model was assumed equivalent to "dead". However, one might want to discriminate between coral recovery and mortality. Yet, again, adding processes implies an increase in the number of parameters to be calibrated. This modifications could lead to an endemic disease dynamics.

@@

% Last nice chapter
The work presented in this thesis has demonstrated the capabilities of modeling tools to inform the management of Florida's Coral Reef. Models allow the evaluation of hypotheses derived from field observations on large-scale systems and can help to deducing the characteristics and underlying mechanisms of complex processes. Furthermore, they can be used in a predictive manner to evaluate the impacts of changing environmental conditions on a given ecosystem. Additionally, the framework developed in this thesis relies on the same tools as already widely used larval transport models and could be combined with them to optimize the design of coral reef protection and restoration campaigns. Given this high potential, further research to improve the model and synergistic work with field an lab studies may well yield results which could help maintaining the biological functions of coral reefs as long as possible through the Anthropocene.
